---
title: "HW2 Review"
author: "Pedro L. Rodríguez"
date: "3/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}

in_path <- "~/Dropbox/NYU/Teaching/Text as Data/homeworks/HW2/data/"
library(dplyr)
library(quanteda)


library(stringr)
library(NLP)
library(tm)
library(RTextTools)
library(utf8)

#libs <- c("dplyr", "jsonlite", "stringr", "foreach", "rjson", "quanteda", "NLP", "tm", "RTextTools", "utf8")
#lapply(libs, library, character.only = TRUE)
#rm(libs)

# Setting WD and seed
in_path <- "/Users/pedrorodriguez/Dropbox/NYU/Teaching/Text as Data/homeworks/HW2/data/"
set.seed(1234)
```

## Part 1

```{r}
# build corpus
df <- tibble(
mystery = c("immigration voter aliens help economy"),
republican1 = c("immigration aliens wall emergency country"),
republican2 = c("voter economy president growth security"),
republican3 = c("healthcare cost socialism unfair help"),
democrat1 = c("immigration country diversity help economy"),   
democrat2 = c("healthcare universal preconditions unfair help"),  
democrat3 = c("economy inequality opportunity voter help"), 
democrat4 = c("abortion choice right women help"))

# create document feature matrix
dfm <- dfm(t(df))
mat <- as.matrix(dfm)
rownames(mat) <- names(df)
mat

# aggregate by author
agg_mat <- rbind(colSums(mat[c(2:4), ]), colSums(mat[c(5:8), ]) )
rownames(agg_mat) <- c("republican", "democrat")          
agg_mat

# num words used by each author
num_words <- rowSums(agg_mat)
```

### (a) NB w/o smoothing:

```{r}
# define priors (# docs of party i/total docs)
democrat_prior <- 4/7
republican_prior <- 3/7

# compute likelihoods
democrat_ll <- prod(agg_mat["democrat", c("immigration", "voter", "aliens", "help", "economy")]/num_words["democrat"])

republican_ll <- prod(agg_mat["republican", c("immigration", "voter", "aliens", "help", "economy")]/num_words["republican"])

# compute posteriors
democrat_post <- democrat_prior*democrat_ll
republican_post <- republican_prior*republican_ll 

# log probs
cat("Posterior probabilities: \n\n Democrat Party:", log(democrat_post), "\n Republican Party:", log(republican_post))
```

Without smoothing we predict the mystery email was sent by the Republican party.

### (b) NB with Laplace smoothing:

```{r}
# apply smoothing (add 1 to all entries):
agg_mat_sm <- agg_mat + 1

# num words used by each author
num_words <- rowSums(agg_mat_sm)

# compute likelihoods
democrat_ll <- prod(agg_mat_sm["democrat", c("immigration", "voter", "aliens", "help", "economy")]/num_words["democrat"])

republican_ll <- prod(agg_mat_sm["republican", c("immigration", "voter", "aliens", "help", "economy")]/num_words["republican"])

# compute posteriors
democrat_post <- democrat_prior*democrat_ll
republican_post <- republican_prior*republican_ll 

# log probs
cat("Posterior probabilities: \n\n Democrat Party:", log(democrat_post), "\n Republican Party:", log(republican_post))
```

With smoothing we predict the mystery email was sent by the Democratic party.

## Part 2

### Question 2

### 2(a): Load data and code positive/negative label based on median score.

```{r}
# data source: https://www.kaggle.com/omkarsabnis/yelp-reviews-dataset

# read in data
samp <- read.csv(paste0(in_path, "yelp.csv"), stringsAsFactors = FALSE)

# compute median of star ratings
med <- median(samp$stars)

# label as positive all ratings above the median
samp$positive <- as.numeric(samp$stars > med)

# inspect balance of labels (note it is quite unbalanced)
prop.table(table(samp$positive))
```

### 2(b): Anchor texts

```{r}
# identify anchor texts
samp <- samp %>% mutate(anchor = ifelse(stars == 5, "positive", ifelse(stars == 1, "negative", "neutral")))

# compute proportions
prop.table(table(samp$anchor))
```

### Question 3: Dictionaries

### 3(a): Hu & Liu sentiment

* Your results may vary depending on pre-processing etc.

```{r}
# read in positive and negative words
pos <- read.table(paste0(in_path, "positive-words.txt"), stringsAsFactors = F)
neg <- read.table(paste0(in_path, "negative-words.txt"), stringsAsFactors = F)

# create dictionary object (a quanteda object)
sentiment_dict <- dictionary(list(pos = pos$V1, neg = neg$V1))

# create document feature matrix with pre-processing options
yelp_dfm <- dfm(samp$text, tolower = TRUE, remove_punct = TRUE, stem = FALSE, dictionary = sentiment_dict)

# calculate net sentiment score
samp$sent <- as.numeric(yelp_dfm[,'pos']) - as.numeric(yelp_dfm[, 'neg'])

# generate binary vector (1 = positive, 0 = negative)
samp$sent_pos <- as.numeric(samp$sent >= 0)

# percent positive
prop.table(table(samp$sent_pos))
```

### 3(b): Histogram

```{r}
hist(samp$sent)
mean(samp$sent_pos)*100
```

### 3(c): Confusion matrix

```{r}
# confusion matrix
confusion_mat <- table(samp$positive, samp$sent_pos)

# baseline accuracy
baseline_accuracy <- max(prop.table(table(samp$positive)))

# accuracy:
dict_acc <- sum(diag(confusion_mat))/sum(confusion_mat) # (TP + TN) / (TP + FP + TN + FN)

# recall:
dict_recall <- confusion_mat[2,2]/sum(confusion_mat[2,]) # TP / (TP + FN)

# precision:
dict_precision <- confusion_mat[2,2]/sum(confusion_mat[,2]) # TP / (TP + FP)

# F1 score:
dict_f1 <- 2*dict_precision*dict_recall/(dict_precision + dict_recall)

# print
cat(
"Baseline Accuracy: ", baseline_accuracy, "\n",
"Accuracy:",  dict_acc, "\n",
"Recall:",  dict_recall, "\n",
"Precision:",  dict_precision, "\n",
"F1-score:", dict_f1
)

```

### 3(d)

```{r}
samp$rank_scores <- rank(samp$stars, ties.method = "average")  

samp$rank_sent <- rank(samp$sent, ties.method = "average")  

rank_sum_dict <- sum(abs(samp$rank_sent - samp$rank_scores))

rank_sum_dict
```

### Question 4: Naive Bayes

* Your results may vary depending on pre-processing etc.

### 4(a)

```{r}
# split sample into training & test sets
set.seed(1984L)
ids <- 1:nrow(samp)
ids_train <- sample(ids, ceiling(0.8*length(ids)), replace = FALSE)
ids_test <- ids[-ids_train]
train_set <- samp[ids_train,]
test_set <- samp[ids_test,]

# get dfm for each set
train_dfm <- dfm(train_set$text, stem = TRUE, remove_punct = TRUE, remove = stopwords("english"))
test_dfm <- dfm(test_set$text, stem = TRUE, remove_punct = TRUE, remove = stopwords("english"))

# match test set dfm to train set dfm features
test_dfm <- dfm_match(test_dfm, features = featnames(train_dfm))

# train model on the training set
nb_model <- textmodel_nb(train_dfm, train_set$positive, smooth = 0, prior = "uniform")

# evaluate on test set
predicted_sentiment <- predict(nb_model, newdata = test_dfm)

# get confusion matrix
cmat <- table(test_set$positive, predicted_sentiment)
nb_acc <- sum(diag(cmat))/sum(cmat) # accuracy = (TP + TN) / (TP + FP + TN + FN)
nb_recall <- cmat[2,2]/sum(cmat[2,]) # recall = TP / (TP + FN)
nb_precision <- cmat[2,2]/sum(cmat[,2]) # precision = TP / (TP + FP)
nb_f1 <- 2*(nb_recall*nb_precision)/(nb_recall + nb_precision)

# print
cat(
"Baseline Accuracy: ", baseline_accuracy, "\n",
"Accuracy:",  nb_acc, "\n",
"Recall:",  nb_recall, "\n",
"Precision:",  nb_precision, "\n",
"F1-score:", nb_f1
)
```

### 4(b)

```{r}
# train model on the training set using docfreq as prior
nb_model <- textmodel_nb(train_dfm, train_set$positive, smooth = 1, prior = "docfreq")

# evaluate on test set
predicted_sentiment <- predict(nb_model, newdata = test_dfm)

# get confusion matrix
cmat <- table(test_set$positive, predicted_sentiment)
nb_acc <- sum(diag(cmat))/sum(cmat) # accuracy = (TP + TN) / (TP + FP + TN + FN)
nb_recall <- cmat[2,2]/sum(cmat[2,]) # recall = TP / (TP + FN)
nb_precision <- cmat[2,2]/sum(cmat[,2]) # precision = TP / (TP + FP)
nb_f1 <- 2*(nb_recall*nb_precision)/(nb_recall + nb_precision)

# print
cat(
"Baseline Accuracy: ", baseline_accuracy, "\n",
"Accuracy:",  nb_acc, "\n",
"Recall:",  nb_recall, "\n",
"Precision:",  nb_precision, "\n",
"F1-score:", nb_f1
)
```

### Question 5: Wordscores

* Your results may vary depending on pre-processing etc.

### 5(a)

```{r}
# get train (anchors) and test (non-anchors) sets
samp_anchor <- samp %>% filter(anchor %in% c("positive", "negative"))
samp_non_anchor <- samp %>% filter(!(anchor %in% c("positive", "negative")))

# get train and test set dfms
train_dfm <- dfm(samp_anchor$text, stem = TRUE, remove = stopwords("english"))
test_dfm <- dfm(samp_non_anchor$text, stem = TRUE, remove = stopwords("english"))

# without smoothing
ws_model <- textmodel_wordscores(train_dfm, ifelse(samp_anchor$anchor == "positive", 1, -1))
hist(ws_model$wordscores)
sort(ws_model$wordscores, decreasing = TRUE)[1:10]
sort(ws_model$wordscores, decreasing = FALSE)[1:10]

# with smoothing
ws_model_sm <- textmodel_wordscores(train_dfm, ifelse(samp_anchor$anchor == "positive", 1, -1), smooth = 1)
hist(ws_model_sm$wordscores)

# Extreme words:
sort(ws_model_sm$wordscores, decreasing = TRUE)[1:10]
sort(ws_model_sm$wordscores, decreasing = FALSE)[1:10]
```

### 5(b) RankSum
```{r}
# RankSum statistic (on smoothed version)

# rank according to stars
samp_non_anchor$rank_scores <- rank(samp_non_anchor$stars, ties.method = "average")  

# get predicted sentiment
predicted_sentiment <- predict(ws_model_sm, newdata = test_dfm)

# rank according to predicted sentiment
samp_non_anchor$rank_ws <- rank(predicted_sentiment, ties.method = "average")

# compute rank difference
rank_sum_ws <- sum(abs(samp_non_anchor$rank_ws - samp_non_anchor$rank_scores))

# which did better?
rank_sum_ws < rank_sum_dict
```

## Question 6: SVM

* Your results may vary depending on pre-processing etc.

### 6(b)

```{r, message = FALSE, cache = TRUE}
# linear SVM model

# subset the reviews
SVM_reviews <- samp[1:1000, ]

# true label
true_label <- factor(SVM_reviews$positive)

# create document feature matrix
library(RTextTools)
reviews_matrix <- create_matrix(SVM_reviews$text, 
                                language = "english", 
                                stemWords = FALSE, 
                                removePunctuation = FALSE, 
                                removeStopwords = TRUE)


# data breaks (will help us change the size of the training set)
training_breaks <- lapply(1:9, function(i) as.integer(0.1*i*nrow(SVM_reviews))) %>% unlist()

# train model
#library(pbapply)
break_accuracy_lin <- lapply(training_breaks, function(i){
  # create container
  container <- create_container(reviews_matrix, 
                                true_label, 
                                trainSize = 1:i,
                                testSize = (i + 1):nrow(SVM_reviews), 
                                virgin = FALSE)
  # fit
  cv.svm <- cross_validate(container, nfold = 5, algorithm = 'SVM', kernel = 'linear')
  return(cv.svm)
})

# plot the mean accuracy for each train data size
lapply(break_accuracy_lin, function(x) x[[2]]) %>% unlist() %>% plot()
max_acc_svm_linear <- lapply(break_accuracy_lin, function(x) x[[2]]) %>% unlist() %>% max()
```

### 6(c) 

```{r, cache = TRUE}
# svm model with radial kernel
break_accuracy_lin <- lapply(training_breaks, function(i){
  # create container
  container <- create_container(reviews_matrix, 
                                true_label, 
                                trainSize = 1:i,
                                testSize = (i + 1):nrow(SVM_reviews), 
                                virgin = FALSE)
  # fit
  cv.svm <- cross_validate(container, nfold = 5, algorithm = 'SVM', kernel = 'radial')
  return(cv.svm)
})

# plot the mean accuracy for each train data size
lapply(break_accuracy_lin, function(x) x[[2]]) %>% unlist() %>% plot()
max_acc_svm_radial <- lapply(break_accuracy_lin, function(x) x[[2]]) %>% unlist() %>% max()

# print comparison
cat("Accuracy: \n\n Linear Kernel:", max_acc_svm_linear, "\n\n Radial Kernel:", max_acc_svm_radial)
```

## 7

### 7(a)
```{r, message = FALSE, cache = TRUE}
library(randomForest)
#https://cran.r-project.org/web/packages/randomForest/randomForest.pdf

# subset the reviews
RF_reviews <- samp[1:500,]

# split original sample into training & test sets
set.seed(1984L)
ids <- 1:nrow(RF_reviews)
ids_train <- sample(ids, ceiling(0.8*length(ids)), replace = FALSE)
ids_test <- ids[-ids_train]
train_set <- RF_reviews[ids_train,]
test_set <- RF_reviews[ids_test,]

# get dfm
train_dfm <- dfm(train_set$text, stem = TRUE, remove_punct = TRUE, remove = stopwords("english"))
test_dfm <- dfm(test_set$text, stem = TRUE, remove_punct = TRUE, remove = stopwords("english"))

# match test set dfm to train set dfm features
test_dfm <- dfm_match(test_dfm, features = featnames(train_dfm))
```

### 7(b)
```{r, cache = TRUE}
# fit model
#start_time <- Sys.time()
rf_model <- randomForest(x = as.matrix(train_dfm), y = factor(train_set$positive), ntree = 501, mtry = sqrt(ncol(as.matrix(train_dfm))), importance = TRUE) 
#Sys.time() - start_time
token_importance <- round(importance(rf_model, 2), 2)
head(rownames(token_importance)[order(-token_importance)])

# plot importance
varImpPlot(rf_model, n.var = 10)
```

### 7(c)
```{r, cache = TRUE}
# baseline accuracy
baseline_accuracy <- max(prop.table(table(test_set$positive)))

# use fitted model to predict labels in test set
predictions_rf <- predict(rf_model, as.matrix(test_dfm))


# get confusion matrix
cmat <- table(test_set$positive, predictions_rf)
rf_acc <- sum(diag(cmat))/sum(cmat) # accuracy = (TP + TN) / (TP + FP + TN + FN)
rf_recall <- cmat[2,2]/sum(cmat[2,]) # recall = TP / (TP + FN)
rf_precision <- cmat[2,2]/sum(cmat[,2]) # precision = TP / (TP + FP)
rf_f1 <- 2*(rf_recall*rf_precision)/(rf_recall + rf_precision)

# print
cat(
"Baseline Accuracy: ", baseline_accuracy, "\n",
"Accuracy:",  rf_acc, "\n",
"Recall:",  rf_recall, "\n",
"Precision:",  rf_precision, "\n",
"F1-score:", rf_f1
)
```


### 7(c)
```{r, cache = TRUE}

# mtry = 0.5*default
rf_model_0.5 <- randomForest(x = as.matrix(train_dfm), y = factor(train_set$positive), ntree = 501, mtry = 0.5*sqrt(ncol(as.matrix(train_dfm))), importance = TRUE) 
predictions_rf_0.5 <- predict(rf_model_0.5, as.matrix(test_dfm))
cmat <- table(test_set$positive, predictions_rf_0.5)
rf_acc_0.5 <- sum(diag(cmat))/sum(cmat) # accuracy = (TP + TN) / (TP + FP + TN + FN)

# mtry = 1.5*default
rf_model_1.5 <- randomForest(x = as.matrix(train_dfm), y = factor(train_set$positive), ntree = 501, mtry = 1.5*sqrt(ncol(as.matrix(train_dfm))), importance = TRUE) 
predictions_rf_1.5 <- predict(rf_model_1.5, as.matrix(test_dfm))
cmat <- table(test_set$positive, predictions_rf_1.5)
rf_acc_1.5 <- sum(diag(cmat))/sum(cmat) # accuracy = (TP + TN) / (TP + FP + TN + FN)

# bind accuracy scores
acc_tb <- tibble(mtry_default = rf_acc, mtry_0.5 = rf_acc_0.5, mtry_1.5 = rf_acc_1.5)

# best performing parameter value
cat("highest out-of-sample accuracy achieved by:", names(acc_tb)[which(acc_tb == max(acc_tb))])
```
